---
title: "Mid_741"
author: "Baiheng Chen"
date: "2024-03-20"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(survival)
library(tidyverse)
```

## Read in Data
```{r}
da <- read.csv("combd_covid19.csv")
colnames(da)
```
## Descriptive statistics
Summarize the demographic (sex, age, ethnicity), socioeconomic (IMD), and clinical (BMI and pre-existing comorbidities) variables by sex and overall. Be sure also to calculate the *death rates*.

"sex","agecat", "ethnicity", "imd", "bmicat"

### Check the missing value
```{r}
sum(is.na(da))
```

### categorize by sex
```{r}
da_male <- da[da$sex == "Male",]
da_female <- da[da$sex == "Female",]
```

### define categroy rate calculation function
```{r}
yes_no_rate<- function(df, col){
  # total observation
  total <- dim(df)[1]
  count_no <- length(df[col][df[col] == "No"])
  no_rate <- count_no/total
  cat(count_no)
  cat("\n")
  cat("The \"No\" rate is:", no_rate)
  cat("\n")
  cat(total - count_no)
  cat("\n")
  cat("The \"Yes\" rate is:", 1- no_rate)
  
  
}
```

### death rate
#### death rate overall
```{r}
# number of death 
sum(da$death == "Yes")/sum(da$time)

```

#### death rate on female
```{r}
sum(da_female$death == "Yes")/sum(da_female$time)
```
#### death rate on male
```{r}
sum(da_male$death == "Yes")/sum(da_male$time)
```
### age
```{r}
cat("Overall")
table(da$agecat)/582
cat("Male")
table(da_male$agecat)/329
cat("Female")
table(da_female$agecat)/253
```
### Ethnicity
```{r}
cat("Female")
table(da_female$ethnicity)/253
cat("Male")
table(da_male$ethnicity)/329
cat("Overall")
table(da$ethnicity)/582

```

### IMD
```{r}
cat("Female")
table(da_female$imd)/253
cat("Male")
table(da_male$imd)/329
cat("Overall")
table(da$imd)/582
```

### BMI
```{r}
cat("Female")
table(da_female$bmicat)/253
cat("Male")
table(da_male$bmicat)/329
cat("Overall")
table(da$bmicat)/582
```

### diabetes1
```{r}
cat("Female")
cat("\n")
yes_no_rate(da_female, "diabetes1")
cat("\n")
cat("Male")
cat("\n")

yes_no_rate(da_male, "diabetes1")
cat("\n")
cat("Overall")
cat("\n")

yes_no_rate(da, "diabetes1")
```



### diabetes2
```{r}
cat("Female")
cat("\n")
yes_no_rate(da_female, "diabetes2")
cat("\n")
cat("Male")
cat("\n")

yes_no_rate(da_male, "diabetes2")
cat("\n")
cat("Overall")
cat("\n")

yes_no_rate(da, "diabetes2")
```



### hypertension
```{r}
cat("Female")
cat("\n")
yes_no_rate(da_female, "hypertension")
cat("\n")
cat("Male")
cat("\n")

yes_no_rate(da_male, "hypertension")
cat("\n")
cat("Overall")
cat("\n")

yes_no_rate(da, "hypertension")
```



### cvd
```{r}
cat("Female")
cat("\n")
yes_no_rate(da_female, "cvd")
cat("\n")
cat("Male")
cat("\n")

yes_no_rate(da_male, "cvd")
cat("\n")
cat("Overall")
cat("\n")

yes_no_rate(da, "cvd")
```



### asthma
```{r}
cat("Female")
cat("\n")
yes_no_rate(da_female, "asthma")
cat("\n")
cat("Male")
cat("\n")

yes_no_rate(da_male, "asthma")
cat("\n")
cat("Overall")
cat("\n")

yes_no_rate(da, "asthma")
```



### copd
```{r}
cat("Female")
cat("\n")
yes_no_rate(da_female, "copd")
cat("\n")
cat("Male")
cat("\n")

yes_no_rate(da_male, "copd")
cat("\n")
cat("Overall")
cat("\n")

yes_no_rate(da, "copd")
```



### cancer
```{r}
cat("Female")
cat("\n")
yes_no_rate(da_female, "cancer")
cat("\n")
cat("Male")
cat("\n")

yes_no_rate(da_male, "cancer")
cat("\n")
cat("Overall")
cat("\n")

yes_no_rate(da, "cancer")
```



### renal
```{r}
cat("Female")
cat("\n")
yes_no_rate(da_female, "renal")
cat("\n")
cat("Male")
cat("\n")

yes_no_rate(da_male, "renal")
cat("\n")
cat("Overall")
cat("\n")

yes_no_rate(da, "renal")
```





## 2.Graphical Analysis

### data transformation
#### combination:
a. type I and II diabetes
b. hypertension and CVD
c. asthma and COPD
```{r}
da_status <- da %>% 
  mutate(status = case_when(death == "No" ~ 0,
                            TRUE~ 1),
         diabetic = case_when(diabetes1 == "No" & diabetes2 == "No" ~ "No", TRUE ~ "Yes"),
         # cardiovascular disease
         card = case_when(hypertension == "No" & cvd == "No" ~ "No",
                          TRUE ~ "Yes"),
         # pulmonary
         pulm = case_when(asthma == "No" & copd == "No" ~ "No",
                          TRUE ~ "Yes"),
         sex = as.factor(sex),
         agecat = as.factor(agecat),
         ethnicity = as.factor(ethnicity),
         imd = as.factor(imd),
         bmicat = as.factor(bmicat),
         diabetic = as.factor(diabetic),
         card = as.factor(card),
         pulm = as.factor(pulm),
         cancer = as.factor(cancer),
         renal = as.factor(renal))
```


### sex
```{r}
png(file = "Sex.png", width = 480, height = 480)
surv_obj_sex <- survfit(Surv(time, status) ~ sex, data = da_status, conf.type = "log-log")
plot(surv_obj_sex, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "Sex", col = c("pink", "skyblue"), lty = 1:2)

legend("topright", legend = levels(da_status$sex), col = c("pink", "skyblue"), lty = 1:2)
dev.off()
```

### age
```{r}
png(file = "Age.png", width = 480, height = 480)
surv_obj_age <- survfit(Surv(time, status) ~ agecat, data = da_status, conf.type = "log-log")
plot(surv_obj_age, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "Age group", col = c("pink", "skyblue", "tomato", "chartreuse","deepskyblue"), lty = 1:5)

legend("bottomleft", legend = levels(da_status$agecat), col = c("pink", "skyblue", "tomato", "chartreuse","deepskyblue"), lty = 1:5)
dev.off()
```

### ethnicity

```{r}
png(file = "Ethnicity.png", width = 480, height = 480)

surv_obj_eth <- survfit(Surv(time, status) ~ ethnicity, data = da_status, conf.type = "log-log")
plot(surv_obj_eth, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "Ethnicity", col = c("pink", "skyblue", "chartreuse"), lty = 1:3)

legend("bottomleft", legend = levels(da_status$ethnicity), col = c("pink", "skyblue", "chartreuse"), lty = 1:3)
dev.off()
```

### imd
```{r}
png(file = "IMD.png", width = 480, height = 480)

surv_obj_imd <- survfit(Surv(time, status) ~ imd, data = da_status, conf.type = "log-log")
plot(surv_obj_imd, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = " The English Indices of Multiple Deprivation", col = c("pink", "chartreuse", "tomato","deepskyblue"), lty = 1:4)

legend("bottomleft", legend = levels(da_status$imd), col = c("pink", "chartreuse", "tomato", "deepskyblue"), lty = 1:4)
dev.off()
```

### bmicat
```{r}
# png(file = "BMI.png", width = 480, height = 480)

surv_obj_bmi <- survfit(Surv(time, status) ~ bmicat, data = da_status, conf.type = "log-log")
plot(surv_obj_bmi, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "BMI", col = c("pink", "skyblue", "chartreuse"), lty = 1:3)

legend("bottomleft", legend = levels(da_status$bmicat), col = c("pink", "skyblue", "chartreuse"), lty = 1:3)
# dev.off()
```

### diabetic
```{r}
png(file = "Diabete.png", width = 480, height = 480)

surv_obj_dia <- survfit(Surv(time, status) ~ diabetic, data = da_status, conf.type = "log-log")
plot(surv_obj_dia, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "Type I & Type II Diabetes", col = c("pink", "skyblue"), lty = 1:2)

legend("bottomleft",legend = levels(da_status$diabetic), col = c("pink", "skyblue"), lty = 1:2)
dev.off()
```



### card
```{r}
png(file = "Card.png", width = 480, height = 480)

surv_obj_card <- survfit(Surv(time, status) ~ card, data = da_status, conf.type = "log-log")
plot(surv_obj_card, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "Cardiovascular Disease", col = c("pink", "skyblue"), lty = 1:2)

legend("bottomleft", legend = levels(da_status$card), col = c("pink", "skyblue"), lty = 1:2)
dev.off()
```

### pulm
```{r}
png(file = "Pulm.png", width = 480, height = 480)

surv_obj_pulm <- survfit(Surv(time, status) ~ pulm, data = da_status, conf.type = "log-log")
plot(surv_obj_pulm, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "Pulmonary Disease", col = c("pink", "skyblue"), lty = 1:2)

legend("bottomleft", legend = levels(da_status$pulm), col = c("pink", "skyblue"), lty = 1:2)
dev.off()
```

### cancer
```{r}
png(file = "Cancer.png", width = 480, height = 480)

surv_obj_cancer <- survfit(Surv(time, status) ~ cancer, data = da_status, conf.type = "log-log")
plot(surv_obj_cancer, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "cancer", col = c("pink", "skyblue"), lty = 1:2)

legend("bottomleft",legend = levels(da_status$cancer), col = c("pink", "skyblue"), lty = 1:2)
dev.off()
```

### renal
```{r}
png(file = "Renal.png", width = 480, height = 480)

surv_obj_renal <- survfit(Surv(time, status) ~ renal, data = da_status, conf.type = "log-log")
plot(surv_obj_renal, ylim = c(0, 1), xlim = c(0, 30), lwd = 2, 
     xlab = "Time (days)", ylab = "Covid-19 Survival Probabilities", 
     main = "Renal disease", col = c("pink", "skyblue"), lty = 1:2)

legend("bottomleft", legend = levels(da_status$renal), col = c("pink", "skyblue"), lty = 1:2)
dev.off()
```




## Multiple regression analysis
```{r}
da_status <- da %>% 
  mutate(status = case_when(death == "No" ~ 0,
                            TRUE~ 1),
         diabetic = case_when(diabetes1 == "No" & diabetes2 == "No" ~ "No", TRUE ~ "Yes"),
         # cardiovascular disease
         card = case_when(hypertension == "No" & cvd == "No" ~ "No",
                          TRUE ~ "Yes"),
         # pulmonary
         pulm = case_when(asthma == "No" & copd == "No" ~ "No",
                          TRUE ~ "Yes"),
         sex = as.factor(sex),
         agecat = as.factor(agecat),
         ethnicity = as.factor(ethnicity),
         imd = as.factor(imd),
         bmicat = as.factor(bmicat),
         diabetic = as.factor(diabetic),
         card = as.factor(card),
         pulm = as.factor(pulm),
         cancer = as.factor(cancer),
         renal = as.factor(renal))
```

### first coxph model

From the plot in Step 2, the survival curve of Pulm is almost the same on "Yes" level and "No" level, so the pulm disease is dropped.
```{r}
obj_1 <- coxph(Surv(time, status) ~ diabetic + card + cancer + renal + sex + agecat + ethnicity + imd, data = da_status)
summary(obj_1)
```

### rescaled Schoelfeld residuals

```{r}
sch <- cox.zph(obj_1) 
sch$table # chi-square tests
```
extremely small p for ethnicity

```{r}
obj_2 <- coxph(Surv(time, status) ~ diabetic + card + cancer + renal + sex + agecat + strata(ethnicity) + imd, data = da_status)
summary(obj_2)
sch <- cox.zph(obj_2) 
sch$table
```

```{r}
anova(obj_2, test="Chisq")
```

### New model
```{r}
da_status <- da_status %>% 
  select(-diabetes1, -diabetes2, -bmicat, -hypertension, -cvd, -asthma, -copd, -pulm)

colnames(da_status)
```

```{r}
da_status <- da_status %>%
  mutate(number = str_count(cancer, "Yes")+str_count(renal, "Yes")+str_count(diabetic, "Yes")+str_count(card, "Yes")) %>% 
  select(-diabetic, -cancer, -card, -renal) 
```
```{r}
obj_3 <- coxph(Surv(time, status) ~ number + sex + agecat + strata(ethnicity) + imd, data = da_status)
summary(obj_3)
```

